Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-04-15 text/back_end_KS.qmd
results chicago

Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity. @fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150). For example, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% had parents who attended some college but did not complete their degree, and 67% were not first-generation college students. However, @fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed their degrees (29% and 28%, respectively). In contrast, Geomarkets such as Chain of Lakes (72%), North Shore (88%), Evanston and Skokie (73%) contribute a disproportionately larger share of White prospects who are not first-generation college students.

@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across all other racial/ethnic categories. The City of Chicago consistently contributed a disproportionately larger share of first-generation college students who parents did not attend college across all racial/ethnic categories. However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects. While Asian, Black, and Hispanic prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.

>>>>

Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity. @fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150) orders. For instance, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% had parents who attended some college but did not complete their degree, and 67% were not first-generation college students. @fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed degrees (29% and 28%, respectively). In contrast, Geomarkets such as Chain of Lakes (72%), North Shore (88%), Evanston and Skokie (73%) contribute a disproportionately larger share of White prospects who are not first-generation college students.

@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories. The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories. However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects. Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares. Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.

These Geomarket contribution patterns become more pronounced as SAT score thresholds increase. For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders. The City of Chicago Geomarket still contributes a disproportionately larger share of first -generation college students - both those whose parents did not attend college and those whose parents attended some college but did not complete a degree - across most racial and ethnic groups. However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%). Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced. All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation college students across both parental education categories to the overall prospect pool.

@fig-chicago-rq2-highSAT-combo shows Geomarket patterns become even more pronounced at the highest SAT score range (1310-1600), largely driven by the relatively fewer number of first-generation college students included within these orders. Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively). Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories.
